---
title: "Replication and Critique of Post-Malthusian Dynamicsin Pre-Industrial Scandinavia"
---

# Purpose

Assessment: 2,500 word essay. Write a referee report that critically dscusses a paper in economic history, focussing on the strengths and weaknesses of the research design. You may modify or extend the analysis. Chose a paper that uses one of the quantitative techniques that we have discuessed in class. Due 5 September 2022.

## Paper

Post-Malthusian Dynamicsin Pre-Industrial Scandinavia by Marc Klemp and Niels Framroze Moller

### What's it about?

Is there a post-malthusian phase in the transition between stagnatio and growth in Scandinavia.

#### What's gonna be hard?

If there's a lot of different data and tests making them digestible is going to be difficult.

Moller and Sharp 2014 "formulate and test a simple dynamic post-Malthusian model. Using an-nual series of real wages, marriage rates, birth rates, and death rates forpre-industrial England, they find evidence consistent with post-Malthusian dynamics emerging as early as the mid-16th century. Likewise, Klemp(2012) formulated and estimated a two-sector Malthusian model on pre-industrial English data and found evidence in favor of Malthusian dynamics exhibiting some post-Malthusian characteristics"


## What to recreate?

Table 1: exerpt of Maddison database. 

```{r}
library(tidyverse)

df <- readxl::read_excel(here::here("data", "mpd2020.xlsx"), sheet = 4, skip = 1)

df <- df %>% 
  pivot_longer(-year) %>% 
  filter(!is.na(value))

library(countrycode)

df <- df %>% 
  mutate(country_name = countrycode(name, "iso3c", "country.name"))

scandi <- c("Sweden", "Denmark", "Norway")

df %>% 
  filter(country_name %in% c(scandi, "United Kingdom")) %>% 
  ggplot(aes(year, value, colour = country_name)) +
  geom_point() +
  geom_line() +
  coord_cartesian(xlim = c(1500, 1900)) +
  scale_y_log10(labels = scales::dollar_format()) 

df_2020 <- df %>% 
  mutate(version = "2020")
```

Table 1: Old Maddison data

```{r}
df <- readxl::read_excel(here::here("data", "md2010.xlsx"), sheet = 4, skip = 2)

df <- df %>% 
  rename(year = ...1) %>% 
  pivot_longer(-year, names_to = "country_name") %>% 
  filter(!is.na(value)) %>% 
  mutate(country_name = case_when(
    country_name == "UK" ~ "United Kingdom",
    TRUE ~ country_name
  ))

scandi <- c("Sweden", "Denmark", "Norway")

df %>% 
  filter(country_name %in% c(scandi, "United Kingdom")) %>% 
  ggplot(aes(year, value, colour = country_name)) +
  geom_point() +
  geom_line() +
  coord_cartesian(xlim = c(1500, 1900)) +
  scale_y_log10(labels = scales::dollar_format()) 

df_2010 <- df %>% 
  mutate(name = countrycode(country_name, "country.name", "iso3c"),
         version = "2010")
```


```{r}
df <- df_2010 %>% 
  bind_rows(df_2020)
```


First critique - which Maddison database to use??

Was Sweden a poor as previously supposed?

```{r}
df %>% 
  filter(country_name %in% c(scandi, "United Kingdom")) %>% 
  ggplot(aes(year, value, colour = version)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ country_name) +
  coord_cartesian(xlim = c(1500, 1900)) +
  scale_y_log10(labels = scales::dollar_format()) 
```


```{r}
df %>% 
  filter(country_name %in% c(scandi, "United Kingdom")) %>% 
  ggplot(aes(year, value, colour = country_name)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ version, nrow = 2) +
  coord_cartesian(xlim = c(1500, 1900)) +
  scale_y_log10(labels = scales::dollar_format()) 
```

## Data

"The estimation is based on yearly observations of real wages, crude birthrates (i.e., the number of births per thousand inhabitants), and crude deathrates (i.e., the number of births per thousand inhabitants) for Denmark,Norway, and Sweden."

### Data Denmark

"The crude demographic rates for Denmark are taken from Mitchell (1998, see their Table A6)"

There is also data for Norway and Sweden in Mitchell. Let's do a comparison.

Pages to extract data from: 112, 113, 116, 120, 123, 127, 130, 133, 134

See ingestion files for data.

```{r}
df <- read_rds("data/mitchell_collected.rds")

df %>%
  # filter(country %in% c("Sweden", "Norway", "Denmark")) %>%
  filter(country %in% c("Denmark")) %>%
  ggplot(aes(year, value, colour = var)) +
  geom_line() +
  facet_wrap(~var, nrow = 3) +
  labs(x = "Year",
       y = "Value of metric per 1000 people",
       title = "Births, deaths and marriages in Denmark",
       subtitle = "Source: Mitchell 1998")
```


### Data Norway

```{r}
# url <- "https://www.ssb.no/a/folketellinger/tabeller/3-13.html"

# library(rvest)
# 
# table <- read_html(url)
# 
# table <- table %>% 
#   html_table()
# 
# table <- table[[3]]
# 
# headers <- table %>% 
#   filter(row_number() == 2)
# 
# headers <- headers %>% 
#   pivot_longer(everything()) %>% 
#   select(value)
# 
# headers <- headers %>% 
#   mutate(english = c("year",
#                      "population",
#                      "born",
#                      "died",
#                      "immigration",
#                      "emmigration",
#                      "married",
#                      "born_out_of_wedlock",
#                      "born_per_1000",
#                      "died_per_1000",
#                      "empty"))
# 
# headers <- headers %>% select(english) %>% pull()
# 
# table <- table %>% 
#   filter(row_number() > 4)
# 
# colnames(table) <- headers
# 
# table <- table %>% 
#   select(-empty) %>% 
#   mutate(across(everything(), parse_number))
# 
# table <- table %>% 
#   mutate(country_name = "Norway")

# table %>% write_rds("data/norway_demographics.rds")
table <- read_rds("data/norway_demographics.rds")
```

### Data Sweden

```{r}
df <- readxl::read_excel("data/Sweden_demographics_lu_madd.xls", range = "A6:F257") %>% janitor::clean_names()

df <- df %>% 
  filter(!is.na(year))

df %>% 
  select(-mid_population) %>% 
  pivot_longer(-year) %>% 
  ggplot(aes(year, value, colour = name)) +
  geom_line()
```

Source: LU MADD - their source: Tabell 2.2 SCB. Demografiska rapporter 1999:2 " Befolkningsutvecklingen under 250 år"

Swedish wages

```{r}
df <- read_rds("data/Joberg_day_wages_sweden.rds")

df %>% 
  ggplot(aes(year, average_price_1)) +
  geom_line()
```


