---
title: "Replication and Critique of Post-Malthusian Dynamicsin Pre-Industrial Scandinavia"
---

# Purpose

Assessment: 2,500 word essay. Write a referee report that critically dscusses a paper in economic history, focussing on the strengths and weaknesses of the research design. You may modify or extend the analysis. Chose a paper that uses one of the quantitative techniques that we have discuessed in class. Due 5 September 2022.

## Paper

Post-Malthusian Dynamicsin Pre-Industrial Scandinavia by Marc Klemp and Niels Framroze Moller

### What's it about?

Is there a post-malthusian phase in the transition between stagnatio and growth in Scandinavia.

#### What's gonna be hard?

If there's a lot of different data and tests making them digestible is going to be difficult.

Moller and Sharp 2014 "formulate and test a simple dynamic post-Malthusian model. Using an-nual series of real wages, marriage rates, birth rates, and death rates forpre-industrial England, they find evidence consistent with post-Malthusian dynamics emerging as early as the mid-16th century. Likewise, Klemp(2012) formulated and estimated a two-sector Malthusian model on pre-industrial English data and found evidence in favor of Malthusian dynamics exhibiting some post-Malthusian characteristics"


## What to recreate?

Table 1: exerpt of Maddison database. 

```{r}
library(tidyverse)

df <- readxl::read_excel(here::here("data", "mpd2020.xlsx"), sheet = 4, skip = 1)

df <- df %>% 
  pivot_longer(-year) %>% 
  filter(!is.na(value))

library(countrycode)

df <- df %>% 
  mutate(country_name = countrycode(name, "iso3c", "country.name"))

scandi <- c("Sweden", "Denmark", "Norway")

df %>% 
  filter(country_name %in% c(scandi, "United Kingdom")) %>% 
  ggplot(aes(year, value, colour = country_name)) +
  geom_point() +
  geom_line() +
  coord_cartesian(xlim = c(1500, 1900)) +
  scale_y_log10(labels = scales::dollar_format()) 

df_2020 <- df %>% 
  mutate(version = "2020")
```

Table 1: Old Maddison data

```{r}
df <- readxl::read_excel(here::here("data", "md2010.xlsx"), sheet = 4, skip = 2)

df <- df %>% 
  rename(year = ...1) %>% 
  pivot_longer(-year, names_to = "country_name") %>% 
  filter(!is.na(value)) %>% 
  mutate(country_name = case_when(
    country_name == "UK" ~ "United Kingdom",
    TRUE ~ country_name
  ))

scandi <- c("Sweden", "Denmark", "Norway")

df %>% 
  filter(country_name %in% c(scandi, "United Kingdom")) %>% 
  ggplot(aes(year, value, colour = country_name)) +
  geom_point() +
  geom_line() +
  coord_cartesian(xlim = c(1500, 1900)) +
  scale_y_log10(labels = scales::dollar_format()) 

df_2010 <- df %>% 
  mutate(name = countrycode(country_name, "country.name", "iso3c"),
         version = "2010")
```


```{r}
df <- df_2010 %>% 
  bind_rows(df_2020)
```


First critique - which Maddison database to use??

Was Sweden a poor as previously supposed?

```{r}
df %>% 
  filter(country_name %in% c(scandi, "United Kingdom")) %>% 
  ggplot(aes(year, value, colour = version)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ country_name) +
  coord_cartesian(xlim = c(1500, 1900)) +
  scale_y_log10(labels = scales::dollar_format()) 
```

