---
title: "Kelly and O Grada (2013)"
format: html
---

## Purpose

Translate the code from session 3 to R.

### Read in data

```{r}
library(tidyverse)
theme_set(theme_light())

df <- readxl::read_excel(here::here("course-content/Class 3/Kelly and O Grada (2013).xlsx"),
                         range = "A1:F256")

# tidy up variable names
df <- df %>% 
  janitor::clean_names()
```

### Changes

```{r}
df <- df %>%
  mutate(
    real_gdp_pc = real_gdp / population,
    real_agricultural_output_pc = real_agricultural_output / population,
    # create logs with prefix ln_
    across(c(crude_death_rate:real_wage, contains("_pc")), log, .names = "ln_{.col}"),
    # create differences with prefix d_
    across(c(contains("ln_")), ~ . - lag(., 1), .names = "d_{.col}"))
```

Now making the lags of the differenced variables:
I don't love this for loop - but I don't know a better way yet. Perhaps using the `recipes` package?

```{r}
for (i in 1:4) {
  message("making d_ for ", i)
  df <- df %>%
    mutate(across(starts_with("d_ln"), ~ lag(., i), .names = paste0("l", i, "_{.col}")))
}
```

### What about using the recipes package?

```{r}
# library(recipes)
# 
# rec <- recipe(d_ln_crude_death_rate ~ ., data = df) %>% 
#   step_log(c(crude_death_rate:real_wage, contains("_pc")), id = rand_id(prefix = "ln_"))
# 
# 
# rec %>% prep() %>% bake(new_data = NULL)
```



```{r}
#| column: page
df %>% 
  skimr::skim()
```

### (Approximately) Reproduce Table 6

What are we doing?

- 3 sets of regressions
- Change in crude death rate is dependent var
- 5 time periods

We should use the many models approach from R4DS because we're doing similar regressions on different sections of the data frame.

```{r}
periods <- tibble(left = c(
  1546,
  1600,
  1650,
  1700,
  1750
)) %>%
  mutate(
    right = case_when(
      left == 1546 ~ 1599,
      TRUE ~ left + 49
    ),
    period = row_number()
  )

periods <- periods %>% 
  group_by(period) %>% 
  mutate(year = paste0(left:right, collapse = ", ")) %>% 
  ungroup() %>% 
  separate_rows(year, sep = ", ", convert = T) %>% 
  dplyr::select(-c(left, right))

df_by_period <- df %>% 
  inner_join(periods)

df_by_period <- df_by_period %>% 
  filter(!is.na(period)) %>%
  group_by(period) %>%
  nest() 
```

Function to do regression

```{r}
reg_model <- function(df, vars){
  df_temp <- df %>% 
    dplyr::select(d_ln_crude_death_rate, vars)
  
  lm(d_ln_crude_death_rate ~ ., data = df_temp)
}
```

### First regression model

```{r}
df_by_period_reg <- df_by_period %>%
  mutate(model = map(data, ~ reg_model(df = .,
                                       vars = 
                                         c(contains("_d_ln_real_wage"), 
                                           contains("d_ln_crude_death_rate")))))
```

What do the regression outputs look like?

```{r}
df_by_period_reg %>% 
  mutate(summary = map(model, summary)) %>% 
  pull(summary)
```

How can we format these better??

```{r}
df_by_period_reg %>% 
  mutate(tidy = map(model, ~ broom::tidy(., conf.int = T))) %>% 
  unnest(tidy) %>% 
  filter(term != "(Intercept)") %>% 
  mutate(term = fct_reorder(term, estimate)) %>% 
  ggplot(aes(estimate, term, colour = factor(period))) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) +
  facet_wrap(~ period) +
  geom_vline(xintercept = 0, lty = 2)
```

```{r}
# df_by_period_reg %>% 
#   mutate(reg = map(model, gtsummary::tbl_regression)) %>% 
#   pull(reg)
```


```{r}
df_by_period_reg %>%
  mutate(augment = map(model, broom::augment)) %>%
  unnest(augment) %>%
  # remove the first four NA years because of the lags (4 lags)
  bind_cols(periods %>% tail(250) %>% dplyr::select(-period)) %>%
  ggplot(aes(year, .fitted, group = period, colour = factor(period))) +
  geom_line(alpha = 1 / 3) +
  geom_smooth(se = FALSE)
```

### Evaluate the residuals

Regression:

dependent variable: first difference of logged crude death rate

independent variables: lags 1 through 4 of first diff of ln crude death rate and first diff of ln real gdp per capita

```{r}
# reg d.LnCrudeDeathRate L(1/4)d.LnCrudeDeathRate L(0/4)d.LnRealGDPpc if tin(1546,1599)
# predict Residuals if tin(1546,1599), res /* generate the residuals as a new variable */ 
# tsline Residuals if tin(1546,1599) /* plot the residuals */
```

```{r}
library(gtsummary)

reg_1 <- df %>%
  filter(between(year, 1546, 1599)) %>%
  dplyr::select(c(
    contains("d_ln_real_gdp_pc"),
    contains("d_ln_crude_death_rate")
  )) %>%
  lm(d_ln_crude_death_rate ~ ., data = .)

reg_1 %>% 
  tbl_regression() %>% 
  add_significance_stars()
```

Plot the residuals

```{r}
reg_1_augmented <- reg_1 %>% 
  broom::augment() %>% 
  bind_cols(periods %>%
    dplyr::select(year) %>%
    slice(5L:54L))

reg_1_augmented %>% 
  ggplot(aes(year, .resid)) +
  geom_line(colour = "red") +
  geom_smooth(method = "lm") +
  geom_hline(yintercept = 0, lty = 2)
```

### Dickey fuller tests

```{r}
library(vars)

adf_test_trend <- ur.df(reg_1_augmented$.resid, type = "trend")

adf_test_trend %>% summary()
```


```{r}
adf_test_none <- ur.df(reg_1_augmented$.resid, type = "none")

adf_test_none %>% summary()
```

```{r}
adf_test_drift <- ur.df(reg_1_augmented$.resid, type = "drift")

adf_test_drift %>% summary()
```

Conclude that the residuals are stationary.

### Breusch–Pagan Test to Check Heteroscedasticity

```{r}
# library(plm)
# 
# df_bp <- df %>%
#   dplyr::filter(dplyr::between(year, 1546, 1599)) %>%
#   dplyr::select(c(
#     contains("d_ln_real_gdp_pc"),
#     contains("d_ln_crude_death_rate")
#   )) %>% na.omit()
# 
# df_bp %>% colnames() %>% paste0(., collapse = " + ")
# 
# plmtest(d_ln_real_gdp_pc ~ l1_d_ln_real_gdp_pc + l2_d_ln_real_gdp_pc + l3_d_ln_real_gdp_pc + l4_d_ln_real_gdp_pc + d_ln_crude_death_rate + l1_d_ln_crude_death_rate + l2_d_ln_crude_death_rate + l3_d_ln_crude_death_rate + l4_d_ln_crude_death_rate, 
#         data = df_bp, 
#         type = "bp")
```

